A Switching Planner for Combined Task and Observation Planning

### 2011-02-08 subarchitectures/planner.sa/trunk/src/latex/AAAI11/evaluation.tex
more small graph changes
few suggestions from moritz

We have implemented our switching continual approach in the MAPSIM
environment~\cite{brenner:nebel:jaamas09},
using \system{dlib-ml}~\cite{king:2009} for belief
revision. Sequential sessions use a modified version of Fast
Downward~\cite{fast-downward}, and DT sessions use our own contingent
procedure. We also implemented a simple dual-mode replanning {\em
baseline} approach. Here, when a switching action is scheduled for
execution the DT session executes a single entropy reduction action,
that directly senses an active assumption -- i.e,. We say a fact can
be directly sensed, if there is a DTPDDL {\em sense} declaration that
mentions the fact in its precondition.  Control is then immediately
returned to a new sequential session.

We evaluate our approaches in robot exploration tasks from home and
office environments. Spatially, these consist of {\em rooms}
(office/kitchen/etc), and an underlying topological map that models
smaller areas of space, called {\em places}, and connectivity between
those. The mobile {\em robot} and {\em visual objects} inhabit the
topological space, the latter indicate the category of space they
inhabit -- e.g., spoons are likely to be in kitchens. By examining
view cones at places for particular objects, the robot is able to: (1)
categorise space at high (room) and low (place) levels, and (2) find
objects for the user, exploiting information about object
co-occurrence and room categories for efficiency. Also, in the
presence of a person, the robot can ask them about the category of the
current room.

We compare switching with the {\em baseline} in several realistic
tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
$|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects, $|$states$|$
$>10^{36}$). We also compare those systems with an approximately
optimal policy computed using Smith's {\sc zmdp} for small 2 room
problems (4-places, 3-objects, $|$states$|$ $\simeq 5000$). For or
experiments we model three sensors: {\em Reliable} sensors have a
$0.1$ probability of a false negative, {\em semi-reliable} have a
chance of $0.3$ of false negative and $0.1$ of false positive , and
finally {\em noisy} sensors with probabilities of $0.5$ and $0.2$
respectively. Each object class is assigned one sensor model, so
e.g. cornflakes may be harder to detect than refrigeratos. To
determine the influence of the sensor model, we performed several
experiments where we only changed the difficulty of the target
object(s). All non-target objects remain unchanged so the planner may
exploit object co-occurances with reliable objects to compensate for
the increased sensor noise.

We evaluate with DT sessions taking $b_0$ admitting between 20 and 100
abstract states with non-zero probability. We run 50 simulations in
each configuration, and have a timeout on each simulation of 30
minutes (1800 seconds)\footnote{All experiments were conducted on a
 2.66GHz Intel Xeon X5355 using one CPU core.}. The continual planning
times are reported in Figure~\ref{fig:results-time}, and the quality
data in Figure~\ref{fig:results-quality}.
%%
For each task, the goal is to find one or several objects and report
their position to a user. Usually there is a non-zero probability that
no plan exists, as the desired object might not be present in the
environment.
% MOG: I didn't do  evaluations for those, so let's remove them
% Item~$f$ reports for tasks requiring indirect sensing,
% where the robot must relocate to a room with a particular category
% (e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
% so is omitted from that reporting.  
Because reward is only allocated
once, on the achievement of the goal,  we find it
intuitive to report average plan costs and the success rates in
problems that admit a solution (i.e., positive reward scaled by a
constant factor).

@@

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-quality}\hfill
  % \vspace{2mm}
  \includegraphics{pomdp-quality}\hfill
  % \includegraphics{pomdp-solvable-quality}\hfill
  \includegraphics{dora2-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora56-quality}\hfill
  % \vspace{2mm}
  % \includegraphics{dora-cat-quality}\hfill
  \caption{Average plan costs and number of successful runs.}
  \label{fig:results-quality}
\end{figure}

We find that if sensing is reliable there is little to be gained
through DT sessions, as the greedy approach of the {\em baseline} is
sufficient. As sensing degrades contingent planning pays off.  Time
spent in DT planning increases steeply as the abstraction $\bstate_0$
becomes more refined.  That refinement seems to be paying off in terms
of the success rate, particularly for tasks $d$ and $e$. For less
refined initial configurations, the increased cost of DT
planning is partially compensated for by a decrease in Fast Downward planning
times. The relatively high success rate irrespective of the level of
refinement in the initial configuration indicates the effectiveness of
using conditional entropy to guide abstraction refinement in our
setting.

>>>>

We have implemented our switching approach in the MAPSIM
environment~\cite{brenner:nebel:jaamas09},
using \system{dlib-ml}~\cite{king:2009} for belief
revision. Sequential sessions use a modified version of Fast
Downward~\cite{fast-downward}, and DT sessions use our own contingent
procedure. We also implemented a simple dual-mode replanning {\em
baseline} approach. Here, when a switching action is scheduled for
execution the DT session applies a single entropy reduction action,
whose execution can provide evidence regarding the truth value of an
active assumption.
%%
%%  -- i.e,. we say a fact can
%% be directly sensed, if there is a DTPDDL {\em sense} declaration that
%% mentions the fact in its precondition. 
%%
Control is then immediately
returned to a new sequential session.

We evaluate our approaches in robot exploration tasks from home and
office environments. Spatially, these consist of {\em rooms}
(office/kitchen/etc), and an underlying topological map over
smaller areas of space, called {\em places}, and connectivity between
those. The mobile {\em robot} and {\em visual objects} inhabit the
topological places. Objects indicate the category of space they
inhabit -- e.g., spoons are likely to be in kitchens. By examining
view cones at places for particular objects, the robot is able to: (1)
categorise space at high (room) and low (place) levels, and (2) find
objects for the user, exploiting information about object
co-occurrence and room categories for efficiency. Also, in the
presence of a person, the robot can ask about the category of the
current room.

We compare switching to the {\em baseline} in several realistic
tasks, with the number of rooms ranging from 3 (12-places, 16-objects,
$|$states$|$ $>10^{21}$) to 6 (26-places, 21-objects, $|$states$|$
$>10^{36}$). We also compare those systems with near optimal policies
computed using Smith's {\sc zmdp} for small 2 room problems (4-places,
3-objects, $|$states$|$ $\simeq 5000$). Our evaluation considers 3
levels of reliability in sensing: {\em reliable} sensors have a $.1$
probability of a false negative, {\em semi-reliable} have a chance of
$0.3$ of false negative and $0.1$ of false positive , and {\em noisy}
sensors with probabilities of $0.5$ and $0.2$ respectively. Each
object class is assigned one sensor model -- e.g. cornflakes may be
harder to detect than refrigerators. We performed several experiments
with different levels of reliability for sensing the target object(s),
while keeping sensing models for non-target objects constant.

\Omit{
All non-target objects remain unchanged so the planner can
exploit object co-occurances using reliably sensed objects to
compensate for the increased sensor noise.
}

We experimented with DT sessions with initial belief-states admitting
between 20 and 100 abstract states with non-zero probability. We run
50 simulations in each configuration, and have a timeout on each
simulation of 30 minutes (1800 seconds)\footnote{All experiments were
conducted on a 2.66GHz Intel Xeon X5355 using one CPU core.}. The
continual planning times are reported in
Figure~\ref{fig:results-time}, and the quality data in
Figure~\ref{fig:results-quality}.
%%
For each task, the goal is to find one or more objects and report
their position to a user. Usually there is a non-zero probability that
no plan exists, as the desired object might not be present in the
environment.
% MOG: I didn't do  evaluations for those, so let's remove them
% Item~$f$ reports for tasks requiring indirect sensing,
% where the robot must relocate to a room with a particular category
% (e.g. kitchen). the {\em baseline} is not able to complete Item~$f$,
% so is omitted from that reporting.  
Because reward is only allocated on the achievement of a goal, we
find it intuitive to report average plan costs and the success rates
in problems that admit a solution (i.e., positive reward scaled by a
constant factor).

@@

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-quality}\hfill
  % \vspace{2mm}
  % \includegraphics{pomdp-solvable-quality}\hfill
  \includegraphics{dora2-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora3-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora4-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora56-quality}\hfill
  \vspace{2mm}
  \includegraphics{pomdp-quality}\hfill
  % \vspace{2mm}
  % \includegraphics{dora-cat-quality}\hfill
  \caption{Average plan costs and number of successful runs.}
  \label{fig:results-quality}
\end{figure}

We find that if sensing is reliable, then little is gained using DT
sessions, as the greedy approach of the {\em baseline} is
sufficient. As sensing degrades DT sessions prove more useful. Here,
time spent on DT planning increases steeply as the abstraction becomes
more refined, which is compensated for by fewer planning sessions
overall. More detailed abstractions lead to a better overall
success rate, particularly for tasks $d$ and $e$.
%% Moreover,
%% for less refined abstractions, the increased cost of DT planning is
%% partially compensated for by a decrease in Fast Downward planning
%% times. 
Speaking to the effectiveness of our entropy heuristic for abstraction
refinement, we witness relatively high success rates irrespective of
the exact level of refinement. Comparing finally to the best {\sc
zmdp} policy, although producing relatively costly plans, the
continual planners performed quite well, especially in terms success
rate. A key source of inefficiency here, is due to sequential sessions
always being optimistic, and refusing to abandon the search.
